### 2018-06-06 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

@@

\cite{Sturm2015} Very good, very important, very close to what I ended up using. Need to look at it further. Maybe hunt down the resutls, they're intersting and interprable, but not by me. The application of artificial neural networks to modelling and generating music is well-studied, e.g., [1–7]. Todd et al. [7] train an RNN with one hidden layer of 8-15 units to reproduce melodies of length 34 notes with quantised time steps. Mozer [6] builds a similar model for melody, but uses an encoding of pitch and time that is pyschoacousti- cally motivated. Eck and Schmidhuber [3] use LSTM to model larger structures than possible with an RNN, and train them on chord progressions and melodies exhibiting 12-bar blues conventions. Eck and Lapamle [2] expands upon this work to model folk tunes, and uses one-hot en- coded input vectors to represent pitches at uniform time delays. Franklin [4] uses a coding scheme similar to [6] to build an RNN with LSTM for pitch and duration, but modelling “jazz-related tasks.” but further removes the ABC fields T: and L:, removes tunes with different meters and/or keys and multiple voices, removes ornaments and gracenotes, transposes all tunes to have the tonic C, (thus giving four modes: major, mixolydian, dorian, and minor), expresses each transcription as a sequence of to- kens (separating pitches, duration, measure bars), and re- moves entries that are not complete transcriptions but in- stead are comments and suggestions, e.g., alternative end- ings. The number of transcriptions in B is 23,636. We use two approaches to modelling and generating ABC. The architecture of our models involves 3 hidden layers with 512 LSTM units each, and a softmax output layer given the distribution over the vocabulary conditioned on the one-hot encoded input. We train each model by back- propogation with one-hot encoded vectors, a mini-batch approach (batch size 50), and with drop out of 0.5. We train our first model using dataset A, with sequences of 50 characters. 3 In the above ABC, examples of characters are “M”, “>” and “:”. The size of the vocabulary is 134. Train- ing makes 100 full passes through the dataset. As an exam- ple, a good model should always predict “:” given an in- put of “M” or “K”, but never given “>”. We train our second model using dataset B with a vocabulary of tokens. Examples of tokens are, “d>”, “K:Amix”, and “|[1”, but not “/” or “:”. Unlike the model above, we train this model using complete tunes: variable length sequences ranging from about 50 to 2000 tokens depending on the tune. 4 We do not split the tunes into unrelated parts, and do not mix parts from different tunes. This preserves the structure of the data. Training makes 100 passes through the dataset. The size of the vocabulary is 137. (with <s> and </s> to mark the start and the end of each tune). To generate ABC from the first model we merely prime it with input text, e.g., “M: 4/4”, and sample any number of times from the distribution at the softmax output. Each generated charamcter produces the next input to the model. For the second model, we prime it with a start symbol <s> and sequentially sample tokens from the softmax outputs until we encounter the end symbol </s> We new look at deeper recurrent architectures (3 hidden layers of 512 LSTM units each), and training on tens of thousands of textual transcriptions of music.
- I CAN"T TOP THIS DATA
We use data retrieved from The Session, 1 an online com- munity of folk music enthusiasts discussing relevant top- ics, and contributing transcriptions of tunes in ABC for- mat.
Regardless, these tran- scriptions are not immediately “session-ready” tunes. Our current work explores refining such models by incorporat- ing corrections made by a domain expert to the output.

\cite{Sturm2016} 
n particular, we build long short-term memory (LSTM) networks having
three hidden layers of 512 LSTM blocks each, and train them using approxi- mately 23,000 music transcriptions expressed with a textual vocabulary (ABC notation). We use this data because it is available, high-level with regards to the music it transcribes, and quite homogeneous with regards to the stylistic conventions of the music (it is crowd-sourced by musicians that play “session” music, e.g., Celtic, Morris, etc.).We take two approaches to training our models: one is character based, in which the system builds a model of joint probabilities of each textual character given the previous 50 characters; the other is “token” based, in which the system computes the joint probability of each token (which can be more than one character) given all previous tokens of a transcription. The result of training is a generative system that outputs transcriptions resembling those in the training material. Our practical aim is to create music transcription models that are useful in particular contexts of music composition, within and outside stylistic conventions particular to the training data. 
Our transcription data comes from a weekly repository of https://thesession. org/,8 an on-line platform for sharing and discussingmusic played in “traditional music sessions” (often Celtic and Morris).
An entry begins with two identifiers, followed by the title, tune type, meter, key, ABC code, date, and contributing user. Contributions vary in detail, with some being quite elaborate, e.g., specifying ornamentation, grace notes, slurs and chords. Most transcriptions are monophonic, but some do specify multiple voices. Many transcriptions have improper ABC formatting, are missing bar lines, have redundant accidentals, miscounted measures, and so on. We create data for training our char-rnn model in the following way.We keep
only five ABC fields (title, meter, key, unit note length, and transcription), and separate each contribution by a blank line. 
This leaves us with a text file having 13,515,723 characters in total, and 47,924 occurrences of T:.10 There are 135 unique characters, e.g., “A”, “:”, and “”, each of which becomes an element of the vocabulary for our char-rnn model.
We create data for training our folk-rnn model in the following way. We
remove title fields and ornaments. We remove all transcriptions that have fewer than 7 measures when considering repetitions (to remove contributions that are not complete transcriptions, but transcriptions of suggested endings, variations, etc.). We remove all transcriptions that have more than one meter or key.11 We transpose all remaining transcriptions (23,636) to a key with root C.
We impose a transcription token vocabulary — each token consists of one or more characters — for the following seven types (with examples in parens): meter (“M:3/4”), key (“K:Cmaj”), measure (“:|” and “|1”), pitch (“C” and “c’”),
grouping (“(3”), duration (“2” and “/2”), and transcription (“<s>” and “<\s>”). 
Our dataset has 4,056,459 tokens, of which 2,816,498 are pitch, 602,673 are du- ration, and 520,290 are measure. A majority of the 23,636 transcriptions consists of 150 tokens or fewer; and 75percent have no more than 190. There are 137 unique tokens, each of which becomes a vocabulary element for our folk-rnn model.

Each LSTM network we build has three hidden layers with 512 LSTM blocks each, and a number of input and output units equal to the number of characters or tokens in its vocabulary. We encode our transcriptions in a local fashion, like in [11, 36], where each element in the vocabulary is mapped to an input and output unit. (This is also called “one-hot encoding”.) The output of each network is a probability distribution over its vocabulary. The total number of parameters in our char-rnn model is 5,585,920; and that in our folk-rnn model is 5,621,722.
All our LSTM networks have the same architecture, but operate over differ- ent vocabularies and are trained differently. One kind we build, which we term char-rnn, operates over a vocabulary of single characters, and is trained on a continuous text file. The second kind we build, folk-rnn, operates over a vocab- ulary of transcription tokens, and is trained on single complete transcriptions. We next discuss our training data, and then the architecture and training of our systems, and finally how we use them to generate new transcriptions.
We initialise the learning rate to 0.002, and ap- ply a decay rate of 0.95 after the first 10 epochs.We build and train our folk-rnn model using our own implementation. This also employs the RMSprop algorithm, but with minibatches of 64 parsed transcriptions each. Since transcriptions in the dataset have different lengths (in number of tokens), we generate minibatches using a bucketing strategy, which places together in a minibatch sequences with approximately the same length, pads them to the maximum length using a “null” token, and then use a masking strategy to ignore null tokens when computing outputs and the loss function. We begin training with a learning rate of 0.003, and a rate decay of 0.97 applied after the 20 first epochs.
For both models, we clip gradients outside [−5, 5] to the limits, and employ
a dropout rate of 0.5 after each LSTM hidden layer. We train each model for 100 epochs in total. We use 95percent of the dataset as training data and 5percent as validation data (the latter for measuring progress in predicting characters or tokens). Through training, our char-rnn model learns a “language model” to produce ABC characters. On the contrary, our folk-rnn model learns a language model in a vocabulary more specific to transcription, i.e., a valid transcription begins with <s>, then a time signature token, a key token, and then a sequence of tokens from 4 types. Our folk-rnn model does not embody the ambiguity of meaning that char-rnn does, e.g., that C can mean a pitch, part of a pitch (C), a letter in a title (A Cup of Tea), or part of a key designation (K:Cmin).
With our trained models, it is a simple matter to have them generate output: we just sample from the probability distribution output by the model over its vocabulary, and use each selected vocabulary element as subsequent input. We can initialise the internal state of each model either randomly, or by inputing a valid “seed” sequence (e.g., beginning with <s>). Repeating the sampling process for N timesteps produces N characters/tokens in addition to the seed sequence.
The role of the composer is clear when we apply our system to create a new
piece of music in Sec. 4.3. Our intention behind seeding the system with the opening two bars of (Fig. 3) is to see how the system responds to an input that does not adhere closely to the stylistic conventions in its training data. Is it able to apply pattern variations even when the input pattern isn’t very close to the learned material? Through our experience, we find the knowledge embedded within the system translates into this different context with the guiding hand of the composer. Within this relatively restricted approach to composition, we find our systems useful for assisting in music material generation that goes in directions we thought little to take.

\cite{Schwarz2006} corpuse based concatanative synthesis method. a large musical corpus is segmented into units and a unit selection algorithm picks the best sequence of units. selection is based on descriptors. The CataRT system is a collection of patches for Max/MSP1 using the FTM, Gabor, and MnM extensions2. The sound and descriptor data can be loaded from SDIF files (see section 4.2) containing MPEG-7 descriptors, for instance, or can be calculated on-the-fly. MODEL CataRT’s model is a multidimensional space of descriptors, pop- ulated by the sound units. The user controls a target point in a lower-dimensional projection of that space with a selection radius around it, and the selection algorithm selects the units closest to the target or within the radius. The actual triggering of the unit is independent of the selection and can happen at any rate. Because displaying and navigating in a high-dimensional space is not practical, the descriptor space is reduced to a 2-dimensional projection according to two selectable descriptors. No concatenation quality is considered, for the moment, and the only transformations applied are a short crossfade to smooth the concatenation and pitch and loudness changes.

@@

ABC \cite{Sturm2015}

>>>>

Уште еден систем за градење на музички мозаици може да се види во \cite{Schwarz2006}, каде авторите имаат изградено корпус од кратки исечоци, секој опишан со одредено множество на дескриптори, наголем дел добиени со математички трансформации на аудио сигналот, пр. брзи фуриеви трансформации, гаворови бранови, хистограми и сл. Системот има алгоритам за избирање на исечоци, кои последователно ги лепи едни за други за време на процесот на генерирање. Алгоритамот е базиран на човечки зададени правила или имитација на постоечки песни, кадешто алгоритамо само ги бира исечоците кои ги задоволуваат критериумите базирани на десктрипторите, без да води сметка за колку добро се сложуваат меѓусебе, и врши благо измазнување на краевиењ помеѓу исечоците.

@@

Во трудот \cite{Sturm2016} авторите користат техники од обработка на природни јазици за генерирање на музика. Караткеристична е употребата на форматот за музика познат како ABC. Форматот е текстуален, наменет примарно за запишување на изворна музика на едноставен и лесно разбирлив начин за луѓето, содржи мелодиска линија запишана со основни тонови и пропратен тескт. Музиката пишана во овој формат е опишана на високо ниво и доста стилски хомогена. Сите овие фактори ги прават форматот и податочното множество одлични за обработка со техники за обработка на природни јазици. Со користење на многу едноставен јазичен модел на ниво на буква и едноставна архитектура од 3 слоја од 512 LSTM неврони, со софтмакс активациски слој авторите добиле многу позитивни резултати. Во трудот прикажуаат и додатно подобрување на системот со повеќе чекори на предпроцесирање на текстот, користејки малку семантичко знаење, го трансформираат текстот од низа букви во низа уникатни и музички значајни токени врз основа на музичката функција на буквите/симболите. Додатно извршиле и процес на селекција и стандардизација на множеството песни за да ги исфрлат песните коишто се: премногу кратки, недволно информативни, конфузни или двосмислени; и го извршиле транскрипција на сите песни во ист клуч. Моделот е трениран со множество од над 23000 песни, во мини-бечови од 50 елементи, во 100 епохи. Генерирањето се врши со итеративна постапка, започнувајќи со посебен симбол за старт. Имаат извршено додатен експеримент во кој го стартуваат моделот со подолг извадок од песни од одредени автори за да ги проверат можностите на моделот да генерализира стилски карактеристки, и авторите тврдат дека моделот е способен за тоа.

@@

ABC \cite{Sturm2015}
ABC notation is a shorthand form of musical notation. In basic form it uses the letters A through G to represent the given notes, with other elements used to place added value on these - sharp, flat, the length of the note, key, ornamentation. Later, with computers becoming a major means of communication, others saw the possibilities of using this form of notation as an ASCII code that could facilitate the sharing of music online, also adding a new and simple language for software developers. In this later form it remains a language for notating music using the ASCII character set. The earlier ABC notation was built on, standardized and changed to better fit the keyboard and an ASCII character set by Chris Walshaw, with the help and input of others. Although now re-packaged in this form, the original ease of writing and reading, for memory jogs and for sharing tunes with others on a scrap of paper or a beer coaster remains, a simple and accessible form of music notation, not unlike others, such as tablature and solfège. Originally designed for use with folk and traditional tunes of Western European origin, e.g., English, Irish, Scottish, which are typically single-voice melodies that can be written on a single staff in standard notation, the work of Chris Walshaw and others has opened this up with an increased list of characters and headers in a syntax that can also support metadata for each tune.[1]

ABC notation being ASCII-based, any text editor can be used to edit the music. Even so, there are now many ABC notation software packages available that offer a wide variety of features, including the ability to read and process ABC notation into MIDI files and as standard "dotted" notation. Such software is readily available for most computer systems, including Microsoft Windows, Unix/Linux, Macintosh, Palm OS, and web-based.[2]

Later third-party software packages have provided direct output, bypassing the TeX typesetter,[3] and have extended the syntax to support lyrics aligned with notes,[4] multi-voice and multi-staff notation,[5] tablature,[6] and MIDI.[7]
